Short-time simulation of open collective quantum spin systems

### 2018-11-29 soc_squeezing/paper/collective_spin.tex
added weak decoherence figure caption and made minor modifications to text

\begin{figure}
  \centering
  \includegraphics{coherent.pdf}
  \caption{Spin squeezing of $N=10^4$ spins under coherent evolution,
    as computed using exact methods (solid lines) and the expansion in
    Eq.~\eqref{eq:order_truncation} with $M=40$ (dashed lines).  Solid
    circles mark the times for TAT and TNT after which the latter
    method gives an unphysical result with $\xi^2<0$.  Both methods
    agree exactly in the OAT case for all times shown.}
  \label{fig:coherent}
\end{figure}

@@

For a collective spin system composed of many small spins, the
operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ contain
genuinely $\abs{\v m}$-body operators (for
$\abs{\v m}\equiv\sum_\mu m_\mu$), but also carry lots of ``baggage''
in the form of $k$-body operators with $k<\abs{\v m}$.  To cut down on
this overhead, we try to change our basis for collective spin
operators from $\S_{\v m}$ to the purely-$\abs{\v m}$-body operators
\begin{align}
  \tilde\S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
  &&
  P_{\v j} \equiv \prod_{\mu,a} s_\mu^{\p{j^\mu_a}},
  \label{eq:S_P}
\end{align}
where $\v m \equiv \p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$ with $m_\mu$ the
number of $s_\mu$ operators in each term of $\S_{\v m}$; $\v j$ is an
ordered list of indices for the spins addressed by $P_{\v j}$; and the
sum over $\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for
which all indices in $\v j$ are distinct and
$\abs{\set{j^\mu_a}} = m_\mu$, such that $j^\mu_a$ indexes the $a$-th
spin addressed by a $s_\mu$ operator in $P_{\v j}$.

@@

where $\simeq$ denotes equality up to a re-indexing of spins; we have
introduced explicit dependence on

@@

$\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in

@@

The sums in \eqref{eq:SS_simp} generally involve a large number of
terms; in practice, many of these terms will be equal to zero.  We
can, however, impose additional restrictions on these sums in such a
way as to throw out terms which vanish, and keep only those which do
not.  Remembering that $s$ counts the number of spins which are
addressed by both operators $P_{\v j}$ and $P_{\v k}$ in
\eqref{eq:SS_PP}, the first restriction we can impose comes from
recognizing that for given $\tilde\S_{\v m},\tilde\S_{\v n}$, the
overlap $s$ is bounded as
\begin{align}
  \max\set{0,\sum_\mu\p{m_\mu+n_\mu}-N}
  \le s \le \min\set{N,\sum_\mu\p{m_\mu+n_\mu}}.
\end{align}
In addition, the structure factor $\eta_{\mu\nu\kappa}$ may generally
contain zeros that we can preemptively avoid with restrictions on
$\v r$ and $\v\rho$.  Recalling that $r_{\mu\nu}$ counts the number of
operators addressed by
$s_\mu s_\nu=\sum_\kappa\eta_{\mu\nu\kappa} s_\kappa$, if
$\eta_{\mu\nu\kappa}=0$ for all $\kappa$, i.e.~$s_\mu s_\nu=0$, then
any term with $r_{\mu\nu}>0$ will vanish.  Our second restriction on
the sums in \eqref{eq:SS_simp} is therefore to fix $r_{\mu\nu}=0$ for
all $\mu,\nu$ with $s_\mu s_\nu=0$.  Finally, we can avoid the
nullifying effect of zeros from $\eta_{\mu\nu\kappa}$
$\v\eta_{\v\rho}\propto\p{\eta_{\mu\nu\kappa}}^{\rho^{\mu\nu}_\kappa}$
by fixing $\rho^{\mu\nu}_\kappa=0$ for all $\mu,\nu,\kappa$ with
$\eta_{\mu\nu\kappa}=0$.  These restrictions are sufficient to ensure
that all individual terms in \eqref{eq:SS_simp} are nonzero, although
they do not rule out the possibility of several terms canceling out.

>>>>

\begin{figure}
  \centering
  \includegraphics{coherent.pdf}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
    under coherent evolution, as computed using exact methods (solid
    lines) and the expansion in Eq.~\eqref{eq:order_truncation} with
    $M=40$ (dashed lines).  Solid circles mark the times for TAT and
    TNT after which the latter method gives an unphysical result with
    $\xi^2<0$.  Both methods agree exactly in the OAT case for all
    times shown.}
  \label{fig:coherent}
\end{figure}

\begin{figure}
  \centering
  % \includegraphics{weak_decoherence.pdf}
  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$ with
    decoherence via uncorrelated decay, excitation, and dephasing of
    individual spins at rates $\gamma_-=\gamma_+=\gamma_\z=\chi$.
    Squeezing is computed using an exact simulation (solid line),
    quantum trajectory methods (dots), and the expansion in
    Eq.~\eqref{eq:order_truncation} with $M=40$ (dashed lines).  Solid
    circles mark the times for TAT and TNT after which the latter
    method gives an unphysical result with $\xi^2<0$.}
  \label{fig:weak_decoherence}
\end{figure}

@@

For a collective spin system composed of many small spins, the
operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ contain
$\abs{\v m}$-body operators (for $\abs{\v m}\equiv\sum_\mu m_\mu$),
but also carry lots of ``baggage'' in the form of $k$-body operators
with $k<\abs{\v m}$.  To cut down on this overhead, we try to change
our basis for collective spin operators from $\S_{\v m}$ to the
purely-$\abs{\v m}$-body operators
\begin{align}
  \tilde\S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
  &&
  P_{\v j} \equiv \prod_{\mu,a} s_\mu^{\p{j^\mu_a}},
  \label{eq:S_P}
\end{align}
where $\v m \equiv \p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$ with $m_\mu$ the
number of $s_\mu$ operators in each term of $\S_{\v m}$; $\v j$ is an
ordered list of indices for the spins addressed by $P_{\v j}$; and the
sum over $\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for
which all indices in $\v j$ are distinct and
$\abs{\set{j^\mu_a}} = m_\mu$, such that $j^\mu_a$ indexes the $a$-th
spin addressed by an $s_\mu$ operator in $P_{\v j}$.

@@

where $\simeq$ denotes equality up to a re-indexing of spins and
additional tensor factors of the single-spin identity operator
(i.e.~$\1$); we have introduced explicit dependence on

@@

$\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in the
index list

@@

The sums in \eqref{eq:SS_simp} generally involve a large number of
terms; in practice, many of these terms will be equal to zero.  We
can, however, impose additional restrictions on these sums in such a
way as to throw out terms which vanish, and keep only those which do
not.  Remembering that $s$ counts the number of spins which are
addressed by both operators $P_{\v j}$ and $P_{\v k}$ in
\eqref{eq:SS_PP}, the first restriction we can impose comes from
recognizing that for given $\tilde\S_{\v m},\tilde\S_{\v n}$, the
overlap $s$ is bounded as
\begin{align}
  \max\set{0,\abs{\v m}+\abs{\v n}-N}
  \le s \le \min\set{N,\abs{\v m}+\abs{\v n}}.
\end{align}
In addition, the structure factor $\eta_{\mu\nu\kappa}$ may generally
contain zeros that we can preemptively avoid with restrictions on
$\v r$ and $\v\rho$.  Recalling that $r_{\mu\nu}$ counts the number of
operators addressed by
$s_\mu s_\nu=\sum_\kappa\eta_{\mu\nu\kappa} s_\kappa$, if
$\eta_{\mu\nu\kappa}=0$ for all $\kappa$, i.e.~$s_\mu s_\nu=0$, then
any term with $r_{\mu\nu}>0$ will vanish.  Our second restriction on
the sums in \eqref{eq:SS_simp} is therefore to fix $r_{\mu\nu}=0$ for
all $\mu,\nu$ with $s_\mu s_\nu=0$.  Finally, we can avoid the
nullifying effect of zeros from
$\v\eta_{\v\rho}=\prod_{\mu,\nu,\kappa}
\p{\eta_{\mu\nu\kappa}}^{\rho^{\mu\nu}_\kappa}$ by fixing
$\rho^{\mu\nu}_\kappa=0$ for all $\mu,\nu,\kappa$ with
$\eta_{\mu\nu\kappa}=0$.  These restrictions are sufficient to ensure
that all individual terms in \eqref{eq:SS_simp} are nonzero, although
they do not rule out the possibility of several terms canceling out.
